# AD SNP overlaps with ATAC peaks

Here we compare the number of AD SNPs overlapping ATAC peaks in different cell types and choosing different thresholds of posterior probability.

```{r Setup, message=FALSE, warning=FALSE, echo=FALSE}
library(tidyverse)

root = "/Users/jeremys/work/opentargets/AD_finemap"
outputPath = "/Users/jeremys/work/opentargets/AD_finemap/analysis"
#setwd(outputPath)

ad.annotated.fname = file.path(root, "annotated", "AD.credible_sets.annotated.colocs.txt")
```

```{r LoadData, warning=FALSE, message=FALSE, echo=FALSE}
ad.df = readr::read_tsv(ad.annotated.fname) %>% dplyr::filter(finemap_prob > 1e-5) %>%
  dplyr::mutate(atac_sum = iPSC + microglia + ipsMacrophage + NPC + Neuron + iNeuron + ipsSensNeuron,
                any_atac = ifelse(atac_sum > 0, 1, 0))

ad.gather.df = ad.df %>%
  dplyr::select(locus_name, rsID, finemap_prob, iPSC, microglia, ipsMacrophage, NPC, Neuron, iNeuron, ipsSensNeuron, atac_sum, any_atac) %>%
  tidyr::gather(key = "cell_type", value = "ATAC_overlap", -locus_name, -rsID, -finemap_prob)
```


```{r PlotPeakSumByCellType, warning=FALSE, message=FALSE, echo=FALSE}
thresholds = c(0.001, 0.002, 0.005, 0.01, 0.02, 0.05, 0.1)
atac.peaksum.df = data.frame()

for (thresh in thresholds) {
  newdf = ad.gather.df %>%
    dplyr::group_by(cell_type) %>%
  dplyr::summarise(threshold = thresh, ATAC_overlaps = sum(ATAC_overlap[finemap_prob > thresh]))
  atac.peaksum.df = rbind(atac.peaksum.df, newdf)
}

ggplot(atac.peaksum.df %>% dplyr::filter(cell_type != "atac_sum"),
       aes(x=threshold, y=ATAC_overlaps, col=cell_type, linetype=cell_type)) +
  geom_line(size=1, alpha=0.8) + geom_point(alhpa=0.6) +
  scale_x_log10() +
  theme_bw() + xlab("SNP PPA threshold") + ylab("Number of ATAC overlaps") +
  ggtitle("Number of ATAC overlaps across AD loci")

ggplot(atac.peaksum.df %>% dplyr::filter(cell_type != "atac_sum"),
       aes(x=threshold, y=ATAC_overlaps, col=cell_type, linetype=cell_type)) +
  geom_line(size=1, alpha=0.8) + geom_point(alhpa=0.6) +
  scale_x_log10() +
  theme_bw() + xlab("SNP PPA threshold") + ylab("Number of ATAC overlaps") +
  ggtitle("Number of ATAC overlaps across AD loci") + coord_cartesian(ylim=c(0, 90))

```

Next, let's see how these overlaps are distributed across loci.

```{r PlotPeakSumByLocus, warning=FALSE, message=FALSE, echo=FALSE}
thresholds = c(0.001, 0.002, 0.005, 0.01, 0.02, 0.05, 0.1)

plotOverlapsByThreshold = function(thresholds) {
  atac.peaksum.df = data.frame()
  
  for (thresh in thresholds) {
    newdf = ad.gather.df %>%
      dplyr::filter(cell_type == "atac_sum") %>%
      dplyr::group_by(locus_name) %>%
      dplyr::summarise(threshold = thresh, ATAC_overlaps = sum(ATAC_overlap[finemap_prob > thresh]))
    atac.peaksum.df = rbind(atac.peaksum.df, newdf)
  }
  atac.peaksum.df$threshold = factor(as.character(atac.peaksum.df$threshold), levels = as.character(thresholds))
  p = ggplot(atac.peaksum.df,aes(x=locus_name, y=ATAC_overlaps, fill=threshold)) +
    geom_bar(stat="identity", position="dodge") +
    theme_bw() + xlab("Locus") + ylab("Total number of ATAC overlaps") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle("Number of ATAC overlaps per locus")
  p
}

print( plotOverlapsByThreshold(thresholds = c(0.001, 0.002, 0.005, 0.01, 0.02, 0.05, 0.1)) )
```

In case that's too many bars to see, look at it with fewer thresholds.

```{r PlotPeakSumByLocus2, warning=FALSE, message=FALSE, echo=FALSE}

print( plotOverlapsByThreshold(thresholds = c(0.005, 0.02, 0.1)) )

```

I will also save a table that shows all SNPs in ATAC peaks, per locus, sorted by decreasing finemap probability.
