# ATAC-seq in iPSC-derived NPCs and neurons

An exploration of ATAC-seq data in iPSC-derived neurons. There are 3 main groups of samples: neurons from the BOB cell line differentiated by overexpression of NEUROG2 (aka NGN2); NPCs from classical neuronal diffentiation (~day 12); neurons from classical differentiation (~day 35).

I called peaks using MACS2 with all samples as input, and peaks with FDR < 1%.

```{r Setup, message=FALSE, warning=FALSE, echo=FALSE}
library(dplyr)
library(readr)
library(magrittr)
library(ggplot2)
library(pheatmap)
library(pcaMethods)
library(DESeq2)

root = "/Users/jeremys/work/opentargets/ipsneurons"
outputPath = "/Users/jeremys/work/opentargets/ipsneurons/GRCh38/ATAC/analysis"
#setwd("/Users/jeremys/work/opentargets/ipsneurons/src/atac/")

counts.fname = file.path(outputPath, "ipsneurons.atac.counts.txt.gz")
counts.mainchrs.fname = file.path(outputPath, "ipsneurons.atac.counts.mainchrs.txt.gz")
narrowpeak.fname = file.path(root, "GRCh38/ATAC/peaks/atac_multisample_peaks.narrowPeak")
```

## Basic QC

Looking across samples, the number of reads varies from 50 M to 150 M.
The fraction of mitochondrial reads varies from 36% to 69% (mean 51%).

```{r CheckChrCounts, warning=FALSE, message=FALSE, echo=FALSE}
readtsv = function(fname, ...) { readr::read_tsv(fname, ...) %>% as.data.frame() }
sample.meta = readtsv(file.path(root, "GRCh38/ATAC/meta/ipsneurons.atac.metadata.txt"))
sample.meta$sampleID = tolower(sample.meta$sampleID)
rownames(sample.meta) = sample.meta$sampleID
sample.meta$day = as.factor(sample.meta$day)

chrCounts.df = readtsv(file.path(root, "GRCh38/ATAC/chrcounts.all.txt"))
colnames(chrCounts.df) = tolower(colnames(chrCounts.df))
counttotals = colSums(chrCounts.df[,-1])
counttotals.nochrM = colSums(chrCounts.df %>% dplyr::filter(grepl("^chr[0-9|X|Y]+$", chr)) %>% dplyr::select(-chr))

chrCounts.df = chrCounts.df %>%
  filter(!grepl("_|\\*", chr)) %>%
  tidyr::gather("sampleID", "count", 2:ncol(chrCounts.df)) %>%
  dplyr::left_join(sample.meta %>% dplyr::select(sampleID, Sample_name))

orderedNames = sample.meta[names(counttotals)[order(counttotals)], ]$Sample_name
chrCounts.df$Sample_name <- factor(as.character(chrCounts.df$Sample_name), levels=orderedNames)

# First just plot the total read counts per sample
ggplot(chrCounts.df %>% arrange(-count),
       aes(x=Sample_name, y=count/1e6, fill=chr)) +
  geom_bar(stat = "identity") + theme_bw(12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  xlab("") + ylab("Number of reads (million)") +
  ggtitle("Reads per chromosome") + guides(fill = guide_legend(ncol = 3)) +
  theme(legend.key.size=unit(0.7, "cm"))

# Do the same using the fraction of reads
chrCounts.df$fraction = chrCounts.df$count / counttotals[chrCounts.df$sampleID]
ggplot(chrCounts.df %>% arrange(-count),
       aes(x=Sample_name, y=fraction, fill=chr)) +
  geom_bar(stat = "identity") + theme_bw(12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  xlab("") + ylab("Fraction") +
  ggtitle("Fraction of reads per chromosome") + guides(fill = guide_legend(ncol = 3)) +
  theme(legend.key.size=unit(0.7, "cm"))

sprintf("Mean fraction of reads from chrM: %.2f", mean(chrCounts.df$fraction[chrCounts.df$chr == "chrM"]))
```

Most peaks are short: 55% are < 200 bp.
The fraction of reads mapping to called peaks varies from 37% to 44%.

```{r CheckReadCounts, warning=FALSE, message=FALSE, echo=FALSE}
# Read peak metadata
allpeak.meta = readtsv(narrowpeak.fname, col_names = c("chr", "start", "end", "peakID", "score", "scrap", "foldChange", "log10pval", "log10qval", "summitPos"))

readCounts = function(fname) {
  counts = readtsv(fname)
  # Fix gene names to remove the version number (e.g. ".4" in ENSG00000223972.4)
  gene_ids <- gsub("\\.[\\d]+", "", counts[,1], perl=T)
  counts[,1] = gene_ids
  rownames(counts) = gene_ids
  counts
}

counts = readCounts(counts.fname)
peak.meta = counts[,c(1,2)]
colnames(peak.meta)[1] = "peakID"
peak.meta = peak.meta %>% dplyr::inner_join(allpeak.meta, by="peakID")
rownames(peak.meta) = peak.meta$peakID

ggplot(peak.meta, aes(x=length)) + geom_histogram(bins=50) + theme_bw(12) +
  ggtitle("ATAC-seq peak size") + xlab("peak size (bp)")
ggplot(peak.meta[peak.meta$length<500,], aes(x=length)) + geom_histogram(bins=50) + theme_bw(12) +
    ggtitle("ATAC-seq peak size") + xlab("peak size (bp)")

# Plot the cumulative distribution function of peak length
peak.size.df <- data.frame(length = unique(peak.meta$length),
                           ecdf = ecdf(peak.meta$length)(unique(peak.meta$length)))
ggplot(peak.size.df[peak.size.df$length<1000,], aes(x=length, y=ecdf)) + geom_line() + theme_bw(12) +
    ggtitle("ATAC-seq peak size cumulative distribution") + xlab("peak size (bp)")

# 55% of peaks are less than 200 bp
#q = quantile(peak.meta$length, probs=seq(0.01,1,0.01))

counts = counts[, -c(1,2)]
colnames(counts) = tolower(colnames(counts))
rownames(counts) = peak.meta$peakID

readCounts.df = data.frame(peakCount = colSums(counts), sampleID = colnames(counts), totalCount = counttotals[colnames(counts)])
readCounts.df$peakFraction = readCounts.df$peakCount / readCounts.df$totalCount
sample.meta2 = sample.meta %>% dplyr::inner_join(readCounts.df, by="sampleID")
ggplot(sample.meta2, aes(x=Sample_name, y=peakFraction)) + geom_bar(stat="identity") + theme_bw(12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Fraction of reads in ATAC-seq peaks") + xlab("")
```

The fraction of reads in peaks when including only main chromosome peaks (no chrM) is slightly higher (43% to 57%).

```{r CheckReadCounts.2, warning=FALSE, message=FALSE, echo=FALSE}
counts = readCounts(counts.mainchrs.fname)
peak.meta = counts[,c(1,2)]
colnames(peak.meta)[1] = "peakID"
peak.meta = peak.meta %>% dplyr::inner_join(allpeak.meta, by="peakID")
rownames(peak.meta) = peak.meta$peakID

counts = counts[, -c(1,2)]
colnames(counts) = tolower(colnames(counts))
rownames(counts) = peak.meta$peakID

readCounts.df = data.frame(peakCount = colSums(counts), sampleID = colnames(counts), totalCount = counttotals.nochrM[colnames(counts)])
readCounts.df$peakFraction = readCounts.df$peakCount / readCounts.df$totalCount
sample.meta2 = sample.meta %>% dplyr::inner_join(readCounts.df, by="sampleID")
ggplot(sample.meta2, aes(x=Sample_name, y=peakFraction)) + geom_bar(stat="identity") + theme_bw(12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Fraction of (non-chrM) reads in ATAC-seq peaks") + xlab("")

meanCounts.df = data.frame(peak=rownames(counts), peakMeanCount=rowMeans(counts))
ggplot(meanCounts.df[meanCounts.df$peakMeanCount < 2^12,], aes(x=log2(peakMeanCount))) +
  geom_histogram(bins=50) + theme_bw(12)
```

Most peaks have reads per million per kb below 2.5, and samples cluster as expected based on their cell type.

```{r CheckReadCounts.3, warning=FALSE, message=FALSE, echo=FALSE}
largepeak.counts = counts[peak.meta$length > 200,]
largepeak.meta = peak.meta[peak.meta$length > 200,]

coldata = data.frame(sampleID = colnames(largepeak.counts))
dds = DESeqDataSetFromMatrix(largepeak.counts, coldata, ~0)
colnames(dds) = colnames(largepeak.counts)
vst = varianceStabilizingTransformation(dds)
dds = estimateSizeFactors(dds)
#View(assay(vst))
counts.deseq = counts(dds, normalized=T)
colnames(counts.deseq) = colnames(largepeak.counts)

rpkm = apply(counts.deseq, MARGIN=2, FUN=function(x) x / sum(x)) * 1e6 * 1e3
rpkm = apply(rpkm, MARGIN=2, FUN=function(x) x / largepeak.meta$length)
peakMedianRpkm = apply(rpkm, MARGIN=1, FUN=median)
peakMeanRpkm = apply(rpkm, MARGIN=1, FUN=mean)

#peakMedianRpkm.df = data.frame(peak=names(peakMedianRpkm[peakMedianRpkm < 50]),
#                               rpkm=peakMedianRpkm[peakMedianRpkm < 50])
#ggplot(peakMedianRpkm.df, aes(x=rpkm)) + geom_histogram(bins = 50) + theme_bw(12) +
#  ggtitle("Reads per kb in ATAC-seq peaks (among those >200 bp)")

peakMedianRpkm.df = data.frame(peak=names(peakMedianRpkm[peakMedianRpkm < 20]),
                               rpkm=peakMedianRpkm[peakMedianRpkm < 20])
ggplot(peakMedianRpkm.df, aes(x=rpkm)) + geom_histogram(bins = 50) + theme_bw(12) +
  ggtitle("Reads per kb in ATAC-seq peaks (among those >200 bp)")

vst.vals.tmp = assay(vst)
colnames(vst.vals.tmp) = sample.meta[colnames(assay(vst)),]$Sample_name
cor.pearson = cor(vst.vals.tmp)
pheatmap(cor.pearson, main="Heatmap of ATAC-seq peak count correlations")
```


